---
title: "Data preparation"
format: html
editor: visual
---

## Imports

```{r}
# This clears all objects in "global environment"
rm(list=ls())

# clears the console area
cat("\014")

# Sets R random numbers to start at the same sequence
set.seed(1234)

install.packages("readr")
install.packages("here")

library(readr)
library(here)
```

## Load Dataset

```{r}
Boston <- read_csv(here::here("data","raw","boston.csv"), show_col_types = FALSE)
head(Boston)
```

## Splitting the Dataset

Split the dataset so that one unique proportion is used to train the model ("train dataset") and one unique set to test the model ("test dataset").

Randomise the dataset just in case it happens to be stored in some order that might impact the model building process; this is random sampling.

```{r}
rowsInDataFrame <- nrow(Boston)
randomNumbers <- runif(rowsInDataFrame) # vector of uniform random numbers between [0.0, 1.0]
randomIndexes <- order(randomNumbers) # creates a vector of 506 unique indexes
```

Use the unique random indexes to extract each row in the dataset in a random order:

```{r}
Boston <- Boston[randomIndexes,]

# Using 70% of the dataset to train the model
training_records <- round(nrow(Boston) * 70/100)
```

```{r}
# train and test sets with a single predictor variable
train_set <- Boston[1:training_records, c("lstat", "medv")]
test_set <- Boston[-(1:training_records), c("lstat", "medv")]
```

## Simple Linear Regression Model
